\thetitle

### 2020-11-18 writing/discussion.tex
discussion: why does transient blame 0 boundaries

An apparent flaw in our transient is the high cost of blame.
Whereas \citet{vss-popl-2017} report an average slowdown of 2.5x and
 worst-case of 5.4x on fully-typed benchmarks in Reticulated,
 our configurations frequently exceed a 10 minute timeout with transient blame.
Unfortunately, our high cost is closer to the truth;
 there are at least three broad issues that skew the earlier results.
First, the 2017 implementation of Reticulated fails to insert certain
 soundness checks\footnote{Missing check: \url{https://github.com/mvitousek/reticulated/issues/36}}
 and blame-map updates\footnote{Missing cast: \url{https://github.com/mvitousek/reticulated/issues/43}}
 from the paper.
Second, Reticulated's type checker assigns the dynamic type to local
 variables in the "fully typed" benchmarks---often because the type system
 cannot articulate a more precise type.
Code that ends up with the dynamic type has fewer constraints to check at run-time.
Third, \citet{vss-popl-2017} use relatively small benchmarks.
Four have since been retired from the official Python benchmark suite
 because they are too small, unrealistic, and unstable.\footnote{Release notes: \url{https://pyperformance.readthedocs.io/changelog.html}}
On larger programs, Reticulated suffers from high overhead.
For example, a Reticulated version of our \texttt{sieve} benchmark runs in
 ~40 seconds without blame, and with blame times out after 10 minutes.

>>>>

Surprisingly, Transient blames an empty set of boundaries in XXXX cases.
Such cases should never occur, at least in theory, because an empty set
 means the value has never crossed a boundary---if the value is indeed defined
 in the current module, then we have typed code blaming itself for a typed
 value; thus, the type checker must be unsound.
After careful investigation of these empty blame cases, we found no soundness
 bugs in Typed Racket.
Instead, we found scenarios in which Transient lost track of the proper
 boundaries.
These scenarios suggest ways to improve the blame algorithm from~\citet{vss-popl-2017}:
\begin{itemize}
  \item
    Entries in a blame map must point to several parent entries.
    For example, if the function \texttt{f} receives bad input in the call
    \texttt{(filter f xs)}, then blame should point to both \texttt{filter}
    and the \texttt{xs} list.
  \item
    The construction of blame map entries must be guided by type-like specifications
     instead of relying on syntax.
    Aliasing the built-in \texttt{filter} function should not change the shape
     of the blame map.
  \item
    The initial blame map must reflect the initial type environment.
    Typed Racket trusts that untyped core-library functions, such as \texttt{filter},
     behave correctly, but does not list the assumption in the blame map.
\end{itemize}
\noindent{}Despite these known limitations, we conjecture that these improvements
 to Transient blame-tracking will not affect our overall conclusions.

A second apparent flaw in our Transient semantics is the high cost of blame.
Whereas \citet{vss-popl-2017} report an average slowdown of 2.5x and
 worst-case of 5.4x on fully-typed benchmarks in Reticulated,
 our configurations frequently exceed a 10 minute timeout with transient blame.
Unfortunately, our high cost is closer to the truth;
 there are at least three broad issues that skew the earlier results.
First, the 2017 implementation of Reticulated fails to insert certain
 soundness checks\footnote{Missing check: \url{https://github.com/mvitousek/reticulated/issues/36}}
 and blame-map updates\footnote{Missing cast: \url{https://github.com/mvitousek/reticulated/issues/43}}
 from the paper.
Second, Reticulated's type checker assigns the dynamic type to local
 variables in the "fully typed" benchmarks---often because the type system
 cannot articulate a more precise type.
Code that ends up with the dynamic type has fewer constraints to check at run-time.
Third, \citet{vss-popl-2017} use relatively small benchmarks.
Four have since been retired from the official Python benchmark suite
 because they are too small, unrealistic, and unstable.\footnote{Release notes: \url{https://pyperformance.readthedocs.io/changelog.html}}
On larger programs, Reticulated suffers from high overhead.
For example, a Reticulated version of our \texttt{sieve} benchmark runs in
 ~40 seconds without blame, and with blame times out after 10 minutes.
